Structured Nonlinear Feedback Design with Separable Control-Contraction Metrics

### 2016-07-19 R0V0 (DESKTOP-K3D88KL's conflicted copy 2016-07-16).tex
> Done: Paper finished. Need to review it.

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor ORCID}

@@

\author{Humberto~Stein Shiromoto,~\IEEEmembership{Member,~IEEE,}
        Ian~R. Manchester,~\IEEEmembership{Member,~IEEE,}% <-this % stops a space
\thanks{Both authors are with the Department of Aerospace, Mechanical, and Mechatronic Engineering and The Australian Centre for Field Robotics. Address: The
Rose Street Building J04, The University of Sydney, NSW 2006, Australia. Corresponding author: humberto.shiromoto@ieee.org. ORCID: 0000-0002-0883-0231, }% <-this % stops a space
}

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
In this paper, input-affine nonlinear systems are considered. The problem under consideration is to design structured (for instance, decentralized or distributed) controllers for this class of systems. Although control-Lyapunov functions are the natural candidate to design these feedback laws, in terms of dynamical programming, the search for these functions is non-convex \cite{Rantzer:2001}. On the other hand, control-contraction metrics provide controllers that render pairs of solutions exponentially converge to each other. Moreover, the search for a structured controller and a suitable metric can be formulated in terms of a convex optimization problem. Under block-diagonal dominance conditions, this convex optimization problem can also be distributed. An example illustrate the approach for a network system.
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
Nonlinear Systems, Feedback Design, Contraction Theory, Distributed Control, Network Systems
\end{IEEEkeywords}

@@

Over the years, methods for the design of controllers for dynamical systems has developed different techniques, according to the vector field that describes the system structure. The main concept employed for the control of dynamic systems is the Lyapunov's stability theory \cite{Bhatia1970,Rouche:1977}. In terms of computation, controller design problems are often translated in terms of dynamic programming.

@@

In the case of nonlinear systems, systematic approaches include high-gain  \cite{GrognardSepulchreBastin1999}, backstepping \cite{Krstic1995}, and forwarding \cite{MazencPraly96}. These methods depend on the structure of the vector field and may fail to be applied when appropriate conditions are not met \cite{SteinShiromoto2013a}. In terms of numerical computation, often the techniques employed for feedback synthesis include sum-of-squares (SOS) optimization \cite{Parrilo2003}.

For those systems described by input-affine differential equations,  the existence of a control-Lyapunov function (CLF) satisfying the so-called \emph{small control property} provides a ``universal'' formula for a stabilizing feedback law  \cite{Sontag1983,Sontag1989a}. However, by the time that this paper was submitted, systematic methods for the search of a CLF were known only for particular systems within this class (see, for instance, \cite{MazencMalisoff2006}). In addition to the lack of a general systematic approach, the search for a CLF in terms of numerical optimization leads to a problem that is not necessarily convex nor connected \cite{Rantzer:2001}.

An alternative to CLFs is provided by the so-called control-contraction metrics (CCMs). One of the main advantages of CCMs over CLFs is that the feedback law is obtained by solving a convex optimization problem \cite{Manchester2014a}, due to fact that the \emph{differential / variational} formulation of the system is considered. The obtained controller makes every solution of the closed-loop system exponentially converge to a prescribed trajectory (see a precise definition below). This stability notion is related to contraction theory (an incremental stability property \cite{Angeli2002,Lohmiller1998,Rueffer2013,Sontag2010,Forni2014}) which can be traced back to the work \cite{Lewis1949}.

The drawback of methods such as CCMs, $H_\infty$ optimization and LQR is that, when the system under consideration has several states (large scale), it does not scale well. This is due to the fact that the stability verification of a system with $n$ states may require a function with $n^2$ terms (see \cite{Rantzer2015} for the case of linear systems and Lyapunov functions).

Large-scale systems may appear in the form of interconnection of smaller components, for example, in applications such as traffic networks \cite{CanudasdeWitMorbidiLeonOjedaEtAl2015}, pipe networks \cite{Persis2011a} and sensor networks \cite{Pajic2011}. For such systems, a property often sought is the ability to augment or to reduce the size of the network without computing the controller for all subsystems, i.e., to make the controller scalable. In some cases, this property is satisfied by desnigning controllers with particular structures (see \cite{Siljak1991,Bai2011} for distributed controllers).

In terms of dynamic programming, the problem of design controllers having an arbitrary structure is known to be NP-hard \cite{BlondelTsitsiklis1997,Tanaka2011}, for linear systems. Consequently, for nonlinear systems the problem is very challenging.

In the case of linear systems whose solutions evolve in the positive orthant (positive systems), the problem of design a structured stabilizing feedback law has been addressed in \cite{Tanaka2011}. Also, the authors show that the existence of such a controller is equivalent to the existence of Lyapunov function with a diagonal structure. The importance of this result for large-scale systems is highlighted in \cite{Rantzer2015}, where sparsity is explored to reduce the number of LMIs to be solved.

For nonlinear systems described by a non-fully connected graph, the design of structured controllers could be simplified by choosing structured Lyapunov functions (see \cite{Dirr2015} and references therein). However, the search for a general Lyapunov function is a non-trivial problem by itself \cite{MalisoffMazenc2009}. The case can be even worse, when structural constraints are considered.

For this reason, the structured design of feedback laws for nonlinear systems is investigated with CCMs, in this work. The systems under consideration are continuously differentiable and affine in the input. One of the key elements employed to design these controllers is the concept of a Riemannian metric (see the definition below) which is employed similarly to Lyapunov functions to measure the exponential convergence of pairs of solutions. Employing CCMs, the design of a structured feedback law allowing exponential convergence of solution is obtained in two steps.

The first step consists of an off-line computation. The original system is extended with its \emph{differential / variational} formulation. The exponential convergence is reformulated in terms of stability of the origin for the extended system. This stability problem is written in terms of a convex optimization problem under the prescribed structural constraint. The solution consists of a pair of matrices: the Riemannian metric and the \emph{differential / variational} feedback law.

The second step consists of an on-line integration. The obtained \emph{differential / variational} controller is integrated along geodesics (see definition below).

\todo[inline]{Highlight the difference between this paper and \cite{Manchester2014a} by saying that this one the integration part is also distributed.}

At this point, the three contributions of this work can be highlighted. First, a systematic framework for off-line computation of a structured controller and metric for nonlinear systems is provided. This framework is the formulation of a convex optimization problem whose computation can also be distributed, provided that some conditions are met. Second, the computation of the online integration phase can distributed, due to the structure imposed on the metric. Third, the structure impose on the controller allows to address the scalability issue of CCMs. Note that the proposed method provides an explicit formula for the controller, since it is constructive.

Other works related to analysis under structural constraints to achieve exponential convergence of solutions do exist in the literature. These are mostly related to network systems. For contraction analysis, see \cite{Aminzare2014a} for analysis according to different network topologies and \cite{Andrieu2016} for transversal contraction analysis. This paper contrast with these works mainly in two points: 1) A systematic framework  for structured controller design is proposed; 2) Due the formulation of the stabilization as a convex optimization problem, not only a explicit formula of a structured controller is obtained but also a metric.

\paragraph{Overview.} The motivation and the problem formulation are provided in Section \ref{sec:Problem Formulation and Motivation}. The background needed for this paper is stated in Section \ref{sec:background}. Section \ref{sec:Results} presents the result of this paper. Illustrations of the proposed approach are provided in Section \ref{sec:Illustration}. The proofs of the results are provided in Section \ref{sec:Proof of the Results}. Section \ref{sec:Conclusion} collects final remarks.

@@

The property described in item 2 of Problem \ref{problem formulation} is particularly relevant for the design feedback laws with a prescribed structure (topology) for network systems rendering it universally stabilizable. Consider the network composed of systems described by the following equation.
\begin{equation}\label{eq:example}
	\left\{\begin{array}{rcl}
		\dot{x}_i&=&-x_i-x_i^3+y_i^2 + 0.01\left(x_{i-1}^3 - 2x_i^3 +x_{i+1}^3\right)\\
		\dot{y}_i&=&u_i,
	\end{array}\right.
\end{equation}
Section \ref{sec:Illustration} shows how the approach proposed in this work is employed to design a decentralized controller for the network composed by interconnections of system \eqref{eq:example}.

\subsection{Background}\label{sec:background}

@@

Let $N>0$ be a constant integer, for each index $i\in\mathbb{N}_{[1,N]}$, consider the system given by
\begin{equation}\tag{\ref{eq:example}}
	\left\{\begin{array}{rcl}
		\dot{x}_i&=&-x_i-x_i^3+y_i^2 + 0.01\left(x_{i-1}^3 - 2x_i^3 +x_{i+1}^3\right)\\
		\dot{y}_i&=&u_i,
	\end{array}\right.
\end{equation}
where for convenience $x_0=x_1$ and $x_N=x_{N+1}$.

@@

In the first one, no constraints were imposed on the matrix $Y$. The parser took  seconds to solve the  MI \eqref{eq:MI formulation}. The components of the matrix $W$ and the matrix $Y$ that form a solution to the \eqref{eq:MI formulation} are
\begin{equation*}
		W_a=\mathbin{\mathtt{diag}}\left(
 1.27,1.53\right),\quad W_b=\mathbin{\mathtt{diag}}\left(1.08,1.53\right)
\end{equation*}
and the matrix $Y$ has the structure shown in Equation~\eqref{eq:example:Y:general} with elements
\begin{align*}
 \overline{Y}_{ij}=&-3y_i-0.08x_iy_i\\
 \underline{Y}_{ij}=&-2+0.009x_i-1.2\sum_{k=1}^4 y_k^2+x_k^2.
\end{align*}

@@

Figure \ref{fig:time graph} shows a plot of the time taken by parser to the MI \eqref{eq:MI formulation} for the case of full decentralization, when the number of agents doubled each time.

\begin{figure}[htpb!]
	\centering
	\includegraphics[width=\columnwidth]{./imgs/ParserTime}
	\caption{Time take by the parser to solve the MI \eqref{eq:MI formulation} for the case of full decentralization, when the number of agents double each time. Axis are on log scale.}
	\label{fig:time graph}
\end{figure}

@@

In this work, input-affine nonlinear systems were considered. The considered problem was the design of structured feedback laws. The methodology employed for synthesis was based on control-contraction metrics which provides a controller ensuring exponential convergence of pairs of solutions.

In terms of dynamic programming, employing control-contraction metrics allows one to formulate the search for the controller as a convex optimization problem. This is one of its main advantages in comparison to the search of control-Lyapunov functions, as its dynamic programming formulation leads to a non-convex problem.

The main contribution of this paper is the synthesis of a feedback law with a prescribed structure. More precisely, since the matrix describing the metric has a block-diagonal structure, the on-line integration phase can be done in parallel, i.e.,  each block defines a metric to the corresponding component of the state space where the geodesic is computed by optimization.

The authors plan to extend this work for the synthesis of event-based controllers, and the analysis of disturbances like time delays.

{\itshape Acknowledgments.} The authors thank the anonymous reviewers.

>>>>

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

@@

\author{Humberto~Stein Shiromoto,~\IEEEmembership{Member,~IEEE,}
        Ian~R. Manchester,~\IEEEmembership{Member,~IEEE,}% <-this % stops a space
\thanks{Both authors are with the Department of Aerospace, Mechanical, and Mechatronic Engineering and The Australian Centre for Field Robotics. Address: The
Rose Street Building J04, The University of Sydney, NSW 2006, Australia. Corresponding author: humberto.shiromoto@ieee.org.}% <-this % stops a space
}

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
The abstract goes here.
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
IEEE, IEEEtran, journal, \LaTeX, paper, template.
\end{IEEEkeywords}

@@

Over the years, methods for the design of controllers for dynamical systems has developed different techniques, according to the vector field that describes the system structure. In terms of computation, controller design problem are often translated in terms of dynamic programming

@@

In the case of nonlinear systems, systematic approaches include high-gain  \cite{GrognardSepulchreBastin1999}, backstepping \cite{Krstic1995,SteinShiromoto2013a}, and forwarding \cite{MazencPraly96}. In terms of numerical computation, techniques employed include sum-of-squares (SOS) optimization \cite{Parrilo2003}.

For those system described by input-affine differential equations,  the existence of a control-Lyapunov function (CLF) satisfying the so-called \emph{small control property} provides a ``universal'' formula for a stabilizing feedback law  \cite{Sontag1983,Sontag1989a}. However, by the time that this paper was submitted, systematic methods for the search of a CLF were known only for particular systems within this class (see, for instance, \cite{MazencMalisoff2006}). In addition to the lack of a general systematic approach, the formulation of the CLF search in terms of numerical optimization leads to a problem that is not necessarily convex and not connected, in general \cite{Rantzer:2001}.

An alternative to CLFs is provided by the so-called control-contraction metrics (CCMs). In this methodology the feedback law is the solution of a convex optimization problem \cite{Manchester2014a}. The obtained controller makes every solution of the closed-loop system exponentially converge to a prescribed trajectory (see a precise definition below).

The drawback of methods such as CCMs, $H_\infty$ optimization and LQR is that, when the system under consideration has several states (large scale), it does not scale well. This is due to the fact that the stability verification of a system with $n$ states may require an appropriate function with $n^2$ terms (see \cite{Rantzer2015} for the case of linear systems and Lyapunov functions).

Large-scale systems may appear in the form of interconnection of smaller components, for example, in applications such as traffic networks \cite{CanudasdeWitMorbidiLeonOjedaEtAl2015}, pipe networks \cite{Persis2011a} and sensor networks \cite{Pajic2011}. For such systems, a property often sought is the ability to augment or to reduce the size of the network without computing the controller for all subsystems of the network, i.e., to make the controller scalable. This property can be satisfied with decentralized controllers \cite{Siljak1991}. In terms of dynamic programming, the problem of design controllers having an arbitrary structure is known to be NP-hard \cite{BlondelTsitsiklis1997,Tanaka2011}, for linear systems. Consequently, for nonlinear systems the problem is very challenging.

In the case of linear systems whose solutions evolve in the positive orthant (positive systems), the problem of design a structured stabilizing feedback law has been addressed in \cite{Tanaka2011}. Also, the authors show that the existence of such a controller is equivalent to the existence of Lyapunov function with a diagonal structure. The importance of this result for large-scale systems is highlighted in \cite{Rantzer2015}, where a structured synthesis of a controller is shown.

For nonlinear systems in which some order of initial conditions is preserved along the flow of the vector field (monotone systems),

\paragraph{Overview.} The motivation and background needed for this paper are stated in Section \ref{sec:Background and Motivation}. The problem under consideration is formalized in Section \ref{sec:Problem Formulation}. Section \ref{sec:Results} presents the result of this paper. Illustrations of the proposed approach are provided in Section \ref{sec:Illustration}. The proofs of the results are provided in Section \ref{sec:Proof of the Results}. Section \ref{sec:Conclusion} collects final remarks.

@@

The property described in item 2 of Problem \ref{problem formulation} is particularly relevant for the design feedback laws with a prescribed structure (topology) for network systems rendering it universally stabilizable. Consider the network composed of systems described by the following equation.
\begin{equation}\label{eq:example system:i}
	\dot{x}_i=??
\end{equation}
Section \ref{sec:Illustration} shows how the approach proposed in this work is employed to design a decentralized controller for the network composed by interconnections of system \eqref{eq:example system:i}.

\subsection{Background}

@@

Let $N>0$ be a constant integer, for each index $i\in\mathbb{N}_{[1,N]}$, consider the system given by
\begin{equation}\label{eq:example}
	\left\{\begin{array}{rcl}
		\dot{x}_i&=&-x_i-x_i^3+y_i^2 + 0.01\left(x_{i-1}^3 - 2x_i^3 +x_{i+1}^3\right)\\
		\dot{y}_i&=&u_i,
	\end{array}\right.
\end{equation}
where for convenience $x_0=x_1$ and $x_N=x_{N+1}$.

@@

In the first one, no constraints were imposed on the matrix $Y$. The parser took 97 seconds to solve the  MI \eqref{eq:MI formulation}. The components of the matrix $W$ and the matrix $Y$ that form a solution to the \eqref{eq:MI formulation} are
\begin{equation*}
		W_a=\mathbin{\mathtt{diag}}\left(
 1.27,1.53\right),\quad W_b=\mathbin{\mathtt{diag}}\left(1.08,1.53\right)
\end{equation*}
and the matrix $Y$ has the structure shown in Equation~\eqref{eq:example:Y:general} with elements
\begin{align*}
 \overline{Y}_{ij}=&-3y_i-0.08x_iy_i\\
 \underline{Y}_{ij}=&-2+0.009x_i-1.2\sum_{k=1}^4 y_k^2+x_k^2.
\end{align*}

@@

@@
